### 2020-11-06 written/SC_model/SC_model.tex
SC model written

\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{tikz-qtree}
\usepackage{tikz-qtree-compat}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{floatrow}

\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\oddsidemargin0cm
\topmargin-2cm
\textwidth16.5cm
\textheight23.5cm

\newcommand{\question}[2] {\vspace{.25in} \hrule\vspace{0.5em}
\noindent{\bf #1: #2} \vspace{0.5em}
\hrule \vspace{.10in}}
\renewcommand{\part}[1] {\vspace{.10in} {\bf (#1)}}
\linespread{1.5}

@@

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%

\begin{document}
\medskip

\thispagestyle{plain}
{\Large Interrogating theoretical models of neural computation with deep inference} \\
Sean R. Bittner$^{1}$, Agostina Palmigiano$^{1}$, Alex T. Piet$^{2,3}$, Chunyu A. Duan$^{4}$, Carlos D. Brody$^{2,3,5}$, \\
Kenneth D. Miller$^{1}$, and John P. Cunningham$^{6}$.

{\small
$^{1}$Department of Neuroscience, Columbia University, \\
$^{2}$Princeton Neuroscience Institute, \\
$^{3}$Princeton University, \\
$^{4}$Institute of Neuroscience, Chinese Academy of Sciences, \\
$^{5}$Howard Hughes Medical Institute, \\
$^{6}$Department of Statistics, Columbia University
}

\linenumbers
\section{Abstract}
A cornerstone of theoretical neuroscience is the circuit model: a system of equations that captures a hypothesized neural mechanism.  
Such models are valuable when they give rise to an experimentally observed phenomenon -- whether behavioral or in terms of neural activity -- and thus can offer insights into neural computation.
The operation of these circuits, like all models, critically depends on the choices of model parameters.
Historically, the gold standard has been to analytically derive the relationship between model parameters and computational properties.  
However, this enterprise quickly becomes infeasible as biologically realistic constraints are included into the model increasing its complexity, often resulting in \emph{ad hoc} approaches to understanding the relationship between model and computation.  
We bring recent machine learning techniques -- the use of deep generative models for probabilistic inference -- to bear on this problem, learning distributions of parameters that produce the specified properties of computation.   
Importantly, the techniques we introduce offer a principled means to understand the implications of model parameter choices on computational properties of interest.  
We motivate this methodology with a worked example analyzing sensitivity in the stomatogastric ganglion.  
We then use it to go beyond linear theory of neuron-type input-responsivity in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and attribute error to connectivity properties in recurrent neural networks solving a simple mathematical task. 
More generally, this work suggests a departure from realism vs tractability considerations, towards the use of modern machine learning for sophisticated interrogation of biologically relevant models.

\section{Introduction}
The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing \cite{abbott2008theoretical}.  
A neural computation is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations.
The function of this system is governed by the choice of model parameters, which when configured in a particular way, give rise to a measurable signature of a computation.   
The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about these particular parameter configurations?  
The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, attractor states and phase transitions, and predictions made by the model.

Consider the idealized practice: one carefully designs a model and analytically derives how model parameters govern the computation.  
Seminal examples of this gold standard (which often adopt approaches from statistical physics) include our field's understanding of memory capacity in associative neural networks \cite{hopfield1982neural}, chaos and autocorrelation timescales in random neural networks \cite{sompolinsky1988chaos}, the paradoxical effect \cite{tsodyks1997paradoxical}, and decision making \cite{wong2006recurrent}.
Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.  
This creates an unfavorable tradeoff.  On the one hand, one may tractably analyze systems of equations with unrealistic assumptions (for example symmetry or gaussianity), mathematically formalizing how parameters affect computation in a too-simple model.  
On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.

% now transition to ML...
Of course, this same tradeoff has been confronted in many scientific fields characterized by the need to do inference in complex models.  
In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observed phenomena (in this case the measurable signal of some computation) back to probability distributions quantifying the likely parameter configurations.  
One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data. 
Indeed, these tools have been used to great success in neuroscience as well, in particular for interrogating parameters (sometimes treated as hidden states) in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}. 
These works have used deep neural networks to expand the expressivity and accuracy of statistical models of neural data \cite{paninski2018neural}.

However, these inference tools have not significantly influenced the study of theoretical neuroscience models, for at least three reasons.  
First, at a practical level, the nonlinearities and dynamics of many theoretical models are such that conventional inference tools typically produce a narrow set of insights into these models.  
Indeed, only in the last few years has deep learning research advanced to a point of relevance to this class of problem.
Second, the object of interest from a theoretical model is not typically data itself, but rather a qualitative phenomenon -- inspection of model behavior, or better, a measurable signature of some computation -- an \emph{emergent property} of the model.  
Third, because theoreticians work carefully to construct a model that has biological relevance, such a model as a result often does not fit cleanly into the framing of a statistical model. 
Technically, because many such models stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation, they lack the explicit likelihood and priors central to the probabilistic modeling toolkit.

% now we've constructed the tension that theory models need this and statistical models know how to do this... now connect!
To address these three challenges, we developed an inference methodology -- `emergent property inference' -- which learns a distribution over parameter configurations in a theoretical model.  
This distribution has two critical properties: \emph{(i)} it is chosen such that draws from the distribution (parameter configurations) correspond to systems of equations that give rise to a specified emergent property (a set of constraints); and \emph{(ii)} it is chosen to have maximum entropy given those constraints, such that we identify all likely parameters and can use the distribution to reason about parametric sensitivity and degeneracies \cite{transtrum2015perspective}.  
First, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2016density, papamakarios2017masked}.
Second, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model, which is ultimately the object of interest in theoretical neuroscience.
Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
Third,  because we cannot assume the theoretical model has explicit likelihood on data or the emergent property of interest, we use stochastic gradient techniques in the spirit of likelihood free variational inference \cite{tran2017hierarchical}.    
Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.

Equipped with this methodology, we then investigated three models of current importance in theoretical neuroscience.
These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to abstract cognitive function), and network scale (from four to infinite neurons).
First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.  
The novel scientific insights offered by EPI contextualize and clarify the previous studies exploring these models \cite{gutierrez2013multiple, litwin2016inhibitory, duan2018collicular, mastrogiuseppe2018linking}, and more generally, these results point to the value of deep inference for the interrogation of biologically relevant models.

%-------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}
\subsection{Motivating emergent property inference of theoretical models} \label{results_motivating}

Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference.  
First, one designs or chooses an existing model that, it is hypothesized, captures the computation of interest. 
To ground this process in a well-known example, consider the stomatogastric ganglion (STG) of crustaceans, a small neural circuit which generates multiple rhythmic muscle activation patterns for digestion \cite{marder2002cellular}.
Despite full knowledge of STG connectivity and a precise characterization of its rhythmic pattern generation, biophysical models of the STG have complicated relationships between circuit parameters and neural activity \cite{goldman2001global, prinz2004similar}.
A model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $z = [g_{\text{el}}, g_{\text{synA}}]$.
Specifically, the two fast neurons ($f1$ and $f2$) mutually inhibit one another, and oscillate at a faster frequency than the mutually inhibiting slow neurons ($s1$ and $s2$).  The hub neuron (hub) couples with either the fast or slow population or both.

Second, once the model is selected, one defines the emergent property, the measurable signal of scientific interest.  
To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.  This emergent property is shown in Figure 1A at a frequency of 0.53Hz.

Third, qualitative parameter analysis ensues: since precise mathematical analysis is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.  Subsequently, a qualitative description is formulated to describe the different parameter configurations that lead to the emergent property.  
In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.  Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.

\begin{figure}
\begin{center}
\includegraphics[scale=0.32]{figures/fig1/fig1.pdf}
\end{center}
\caption{\small Emergent property inference (EPI) in the stomatogastric ganglion.  A. For a choice of model (STG) and emergent property (network syncing), emergent property inference (EPI, gray box) learns a distribution of the model parameters $z = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing network syncing.  In the STG model, jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$. Other connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.  Network syncing traces are colored by log probability density of their generating parameters (stars) in the EPI-inferred distribution.  B. The EPI distribution of STG model parameters producing network syncing.  Samples are colored by log probability density.  Distribution contours of emergent property value error are shown at levels of $2.5 \times 10^{-5}$, $5 \times 10^{-5}$, $1 \times 10^{-4}$, $2 \times 10^{-4}$, and $4 \times 10^{-4}$ (dark to light gray).  Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $v_1$ (solid) and $v_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.  Simulated activity is shown for three samples (stars). (Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  $v_1$ is sensitive to network syncing ($p<10^{-4}$), while $v_2$ is not ($p=0.67$) (see Section \ref{methods_STG}).  C. Deep probability distributions map a latent random variable $w$ through a deep neural network with weights and biases $\theta$ to parameters $z = f_\theta(w)$ distributed as $q_\theta(z)$. D. EPI optimization: To learn the EPI distribution $q_\theta(z)$ of model parameters that produce an emergent property, the emergent property statistics $T(x)$ are set in expectation over model parameter samples $z \sim q_\theta(z)$ and model simulations $x \sim p(x \mid z)$ to emergent property values $\mu$.}
 \label{fig:STG}
\end{figure}

Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig. \ref{fig:STG}A gray box) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution shown in Figure \ref{fig:STG}B.  
This distribution -- represented for clarity as samples from the distribution -- is then a scientifically meaningful and mathematically tractable object.  
In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for network syncing (the mode; Figure \ref{fig:STG}B yellow star), and how network syncing decays based on changes away from the mode.
The eigenvectors (of the Hessian of the distribution at the mode) quantitatively formalize the robustness of network syncing (Fig. \ref{fig:STG}B solid ($v_1$) and dashed  ($v_2$) black arrows).
Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$) and degeneracy ($v_2$) agree with error contours (Fig. 1B, contours) and have diminished or preserved network syncing, respectively (Figure 1B inset and activity traces) (see Section \ref{methods_STG}).

\subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}

Emergent property inference (EPI) systematizes the three-step procedure of the previous section.
First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  In the running STG example, the model activity $x = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
\begin{equation} C_m \frac{dx}{dt} = -h(x; z) = - \left[ h_{leak}(x; z) + h_{Ca}(x; z) + h_K(x; z) + h_{hyp}(x; z) + h_{elec}(x; z) + h_{syn}(x; z)\right] 
\end{equation} 
where $C_m$=1nF, and $h_{\text{leak}}$, $h_{Ca}$, $h_K$, $h_{\text{hyp}}$, $h_{\text{elec}}$, and $h_{\text{syn}}$ are the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on $x$ and $z = [g_{\text{el}}, g_{\text{synA}}]$ (see Section \ref{methods_STG}).

Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(x)$, $\omega_{\text{f2}}(x)$, etc. -- is close to an intermediate frequency of 0.53Hz.  
Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(x)$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
\begin{equation}\label{eq:EP}
 \mathbb{E}\left[T(x) \right] ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \omega_{\text{f1}}(x) \\ \vdots \\ (\omega_{\text{f1}}(x) - 0.53)^2 \\ \vdots \end{bmatrix} ~~=~~  
 \begin{bmatrix} 0.53 \\ \vdots \\ 0.025^2 \\ \vdots \end{bmatrix} ~~\triangleq~~ \mu,
 \end{equation}
  which completes the quantification of the emergent property.

Third, we perform emergent property inference: we find a distribution over parameter configurations $z$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP}.  
This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_\theta(z) : \theta \in \Theta \right\}$, defined by a deep generative distribution
of the normalizing flow class \cite{rezende2015variational, dinh2016density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).  
This deep distribution is represented in Figure \ref{fig:STG}C (see Section \ref{methods_EPI}).  
Then, mathematically, we must solve the following optimization program: 
 \begin{equation} \label{eq:EPI}
\begin{split}
&\argmax_{q_\theta \in \mathcal{Q}} H(q_\theta(z)) \\
 &\text{  s.t.  } \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu, \\
\end{split}
\end{equation}
where $T(x), \mu$ are defined as in Equation \ref{eq:EP}, and $p(x|z)$ is the intractable distribution of data from the model, $x$, given that model's parameters $z$ (we access samples from this distribution by running the model forward).   The purpose of each element in this program is detailed in Figure \ref{fig:STG}D.
Finally, we recognize that many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select that which has maximum entropy.
This principle, captured in Equation \ref{eq:EPI} by the primal objective $H$, identifies parameter distributions with minimal assumptions beyond some chosen structure \cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}.
Such a normative principle of maximum entropy, which is also that of Bayesian inference, naturally fits with our scientific objective of reasoning about parametric sensitivity and robustness. The recovered distribution of EPI is as variable as possible along each parametric manifold such that it produces the emergent property.

EPI optimizes the weights and biases $\theta$ of the deep neural network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
The optimization is complete when the sampled models with parameters $z \sim q_\theta$ produce activity consistent with the specified emergent property (Fig. S4).
Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
Further validation of EPI is available in the supplementary materials, where we analyze a simpler model for which ground-truth statements can be made (Section \ref{methods_2DLDS}).

In relation to broader methodology, inspection of the EPI objective reveals a natural relationship to posterior inference.
Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by the emergent property statistics and values, respectively (see Section \ref{methods_VI}).
Equipped with this method, we may examine structure in such inferred distributions or make comparisons between inferred distributions at graded values of emergent property statistics.
We now prove out the value of EPI by using it to investigate and produce novel insights about three prominent models in neuroscience.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comprehensive input-responsivity in a nonlinear sensory system} \label{results_V1}

\begin{figure}
\begin{center}
\includegraphics[scale=0.54]{figures/fig2/fig2.pdf}
\end{center}
\caption{\small Hypothesis generation through EPI in a V1 model.  A. Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons.   Some neuron-types largely do not form synaptic projections to others  (excitatory and inhibitory projections filled and unfilled, respectively).  B. Linear response predictions become inaccurate with greater input strength.  V1 model simulations for input (solid) $h=b$ and (dashed) $h = b + dh$. Stars indicate the linear response prediction.  C. EPI distributions on differential input $dh$ conditioned on differential response $\mathcal{B}(\alpha, y)$. Supporting evidence for the four generated hypotheses are indicated by gray boxes with labels H1, H2, H3, and H4. The linear prediction from two standard deviations away from $y$ (from negative to positive) is overlaid in magenta (very small, near origin). }
 \label{fig:V1_EPI}
\end{figure}

Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}. 
Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig. \ref{fig:V1_EPI}A) \cite{pfeffer2013inhibition}.
Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools. 
Here, we go beyond linear theory by systematically generating and evaluating hypotheses of circuit model function using EPI distributions of neuron-type inputs producing various neuron-type population responses.

Specifically, we consider a four-dimensional circuit model with dynamical state given by the firing rate $x$ of each neuron-type population $x = \left[x_E, x_P , x_S, x_V \right]^\top$. Given a time constant of $\tau = 20$ ms and a power $n = 2$, the dynamics are driven by the rectified and exponentiated sum of recurrent ($Wx$) and external $h$ inputs:

\begin{equation}
\tau \frac{dx}{dt} = -x + [W x+ h]_+^n.
\end{equation}

We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic} (see Section \ref{methods_V1}).
The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top$ and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.  
Throughout subsequent analyses, the baseline input is $b = \left[ 1 ,1,1,1\right]^\top$.

With this model, we are interested in the differential responses of each neuron-type population to changes in input $dh$. 
Initially, we studied the linearized response of the system to input $\frac{dx_{ss}}{dh}$ at the steady state response $x_{ss}$, i.e. a fixed point. 
All analyses of this model consider the steady state response, so we drop the notation $ss$ from here on.
While this linearization accurately predicts differential responses $dx = \left[ dx_{E} , dx_{P} , dx_{S} ,dx_{V} \right]^\top$  for small differential inputs to each population $dh = \left[ 0.1 , 0.1 , 0.1 , 0.1 \right]^\top$ (Fig \ref{fig:V1_EPI}B left), the linearization is a poor predictor in this nonlinear model more generally (Fig. \ref{fig:V1_EPI}B right).  
Currently available approaches to deriving the steady state response of the system are limited.

To get a more comprehensive picture of the input-responsivity of each neuron-type beyond linear theory, we used EPI to learn a distribution of the differential inputs to each population $dh$ that produce an increase of $y$ in the rate of each neuron-type population $\alpha \in \{E, P, S, V \}$.  
We want to know the differential inputs $dh$ that result in a differential steady state $dx_{\alpha}$ (the change in $x_{\alpha}$ when receiving input $h=b + dh$ with respect to the baseline $h = b$) of value $y$ with some small, arbitrarily chosen amount of variance  $0.01^2$.   
These statements amount to the emergent property 
\begin{equation}
\mathcal{B}(\alpha, y) ~~\triangleq~~ 
\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ (dx_{\alpha} - y)^2 \end{bmatrix} ~~=~~ \begin{bmatrix} y \\ 0.01^2 \end{bmatrix}.
\end{equation}
We maintain the notation $\mathcal{B}(\cdot)$ throughout the rest of the study as short hand for emergent property, which represents a different signature of computation in each application.

Using EPI, we inferred the distribution of $dh$ shown in Figure \ref{fig:V1_EPI}C producing $\mathcal{B}(\alpha, y)$.  
Columns correspond to inferred distributions of excitatory ($\alpha=E$, red), parvalbumin ($\alpha=P$, blue), somatostatin ($\alpha=S$, red) and VIP ($\alpha=V$, green) neuron-type response increases, while each row corresponds to increase amounts of $y \in \{0.1, 0.5\}$.  For each pair of parameters, we show the two-dimensional marginal distribution of samples colored by $\log q_\theta(dh \mid \mathcal{B}(\alpha, y))$.  The inferred distributions immediately suggest four hypotheses: \\

{\addtolength{\leftskip}{10 mm}
H1: as is intuitive, each neuron-type's firing rate should be sensitive to that neuron-type's direct input (e.g. Fig. \ref{fig:V1_EPI}C H1 gray boxes indicate low variance in $dh_E$ when $\alpha=E$. Same observation in all inferred distributions); \\
H2: the E- and P-populations should be largely unaffected by input to the V-population (Fig. \ref{fig:V1_EPI}C H2 gray boxes indicate high variance in $dh_V$ when $\alpha \in \{E, P \}$); \\
H3: the S-population should be largely unaffected by input to the P-population  (Fig. \ref{fig:V1_EPI}C H3 gray boxes indicate high variance in $dh_P$ when $\alpha =S$); \\
H4: there should be a nonmonotonic response of the V-population with input to the E-population (Fig. \ref{fig:V1_EPI}C H4 gray boxes indicate that negative $dh_E$ should result in small $dx_{V}$, but positive $dh_E$ should elicit a larger $dx_{V}$);

}

\begin{figure}
\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=4cm}}]{figure}[\FBwidth]
{\caption{Confirming EPI generated hypotheses in V1. A. Differential responses $\delta x_E$ by the E-population to changes in individual input $\delta h_\alpha \hat{u}_\alpha$ away from the mode of the EPI distribution $dh^*$. B-D Same plots for the P-, S-, and V-populations.  Labels H1, H2, H3, and H4 indicate which curves confirm which hypotheses.}\label{fig:V1_HT}}
{\includegraphics[scale=0.6]{figures/fig3/fig3.pdf}}
\end{figure}
We evaluate these hypotheses by taking perturbations in individual neuron-type input $\delta h_\alpha$ away from the modes of the inferred distributions at $y=0.1$
\begin{equation}
dh^* = z^* = \argmax_{z} \log q_\theta(z \mid \mathcal{B}(\alpha, 0.1)).
\end{equation}
Here $\delta x_{\alpha}$ is the change in steady state response of the system with input $h = b + dh^* + \delta h_\alpha \hat{u}_\alpha$ compared to $h = b + dh^*$, where $\hat{u}_\alpha$ is a unit vector in the dimension of $\alpha$. The EPI-generated  hypotheses are confirmed (for details, see Section \ref{methods_V1}):

{\addtolength{\leftskip}{10 mm}
H1: the neuron-type responses are sensitive to their direct inputs (Fig. \ref{fig:V1_HT}A black, \ref{fig:V1_HT}B blue, \ref{fig:V1_HT}C red, \ref{fig:V1_HT}D green); \\
H2: the E- and P-populations are not affected by $\delta h_V$ (Fig. \ref{fig:V1_HT}A green, \ref{fig:V1_HT}B green); \\
H3: the S-population is not affected by $\delta h_P$ (Fig. \ref{fig:V1_HT}C blue); \\
H4: the V-population exhibits a nonmonotonic response to $\delta h_E$ (Fig. \ref{fig:V1_HT}D black), and is in fact the only population to do so (Fig. \ref{fig:V1_HT}A-C black).

}

These hypotheses were in stark contrast to what was available to us via traditional analytical linear prediction (Fig. \ref{fig:V1_EPI}C, magenta, see Section \ref{methods_V1}).

Here, we examined the neuron-type responsivity of this model of V1 with scientifically motivated choice of connectivity $W$.
With EPI, we could just as easily have examined the distribution of such $W$'s consistent with some response characteristics for a fixed input $h$ or another emergent property such as inhibition stabilization.
Most importantly, this analysis is a proof-of-concept demonstrating the valuable ability to condition parameters of interest of a neural circuit model on some chosen emergent property.
To this point, we have shown the utility of EPI on relatively low-level emergent properties like network syncing and differential neuron-type population responses.  
In the remainder of the study, we focus on using EPI to understand models of more abstract cognitive function.

>>>>

% You should title the file with a .tex extension (hw1.tex, for example)
\documentclass[11pt]{article}

\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{tikz-qtree}
\usepackage{tikz-qtree-compat}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usepackage{systeme}
\usepackage{graphicx}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\oddsidemargin0cm
\topmargin-2cm     %I recommend adding these three lines to increase the 
\textwidth16.5cm   %amount of usable space on the page (and save trees)
\textheight23.5cm

\newcommand{\question}[2] {\vspace{.25in} \hrule\vspace{0.5em}
\noindent{\bf #1: #2} \vspace{0.5em}
\hrule \vspace{.10in}}
\renewcommand{\part}[1] {\vspace{.10in} {\bf (#1)}}

\newcommand{\myname}{Sean Bittner}
\newcommand{\myandrew}{srb2201@columbia.edu}
\newcommand{\myhwnum}{12}

@@

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
 %
\pagestyle{fancyplain}
\rhead{\fancyplain{}{\myname\\ \myandrew}}

\begin{document}

\medskip                        % Skip a "medium" amount of space
                                % (latex determines what medium is)
                                % Also try: \bigskip, \littleskip

\thispagestyle{plain}
\begin{center}                  % Center the following lines
{\Large Draft of new SC section} \\
Sean Bittner \\
November 4, 2020 \\
\end{center}

\section{Identifying neural mechanisms of flexible task switching} \label{results_SC}
In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig. \ref{fig:SC}a). 
Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
Duan et al. proposed a model of SC that, like the V1 model analyzed in the previous section, is a four-population dynamical system.  
We analyzed this model, where the neuron-type populations are functionally-defined as the Pro- and Anti-populations in each hemisphere (left (L) and right (R)), their connectivity is parameterized geometrically  (Fig. \ref{fig:SC}B).
The input-output function of this model is chosen such that the population responses $x = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 giving rise to high (1) or low (0) responses at the end of the trial:
\begin{equation}
x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
\end{equation}
where $\epsilon = 0.05$ and $\zeta = 0.5$.  
The dynamics evolve with timescale $\tau=0.09$ via an internal variable $u$ governed by connectivity weights $W$
\begin{equation}
\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
\end{equation}
with gaussian noise of variance $\sigma^2$ = 1.
The input $h$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
Here, we use EPI to determine the changes in network connectivity $z = [ sW_P, sW_A, vW_{PA}, vW_{AP}, dW_{PA}, dW_{AP}, hW_P, hW_A]$ resulting in greater levels of rapid task switching accuracy.

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figs/fig4.pdf}
\end{center}
\caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  
A. Rapid task switching behavioral paradigm (see text). 
B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
C. The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$. Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.  
D. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
E. The correlation of Schur eigenvalue with task performance in each learned EPI distribution.}
\label{fig:SC}
\end{figure}
